Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-09-17 fib2.tex
revised responses to referee comments

\newcommand{\Eref}[1]{Eq.~(\ref{#1})}
\newcommand{\Fref}[1]{Fig.~\ref{#1}}

@@

Concretely, we consider the simulation of quantum error correction in a two-dimensional lattice 
system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum 
computation~\cite{Freedman2002, Nayak2008}. Fibonacci anyons are experimentally motivated as the 
expected excitations of the $\nu=\frac{12}{5}$ fractional quantum Hall 
states~\cite{Slingerland2001}, and can be realized in several spin 
models~\cite{Levin2005, Bonesteel2012, Kapit2013, Palumbo2014} and composite 
heterostructures~\cite{Mong2014}.
Any of these kinds of physical systems could be used to perform universal topological quantum computation \cggb{citations?}. Natural sources of noise from thermal fluctuations or external perturbations will be suppressed by the energy gap but must still be corrected to allow for scalable computation.

We use a flexible phenomenological model of dynamics and thermal 
noise to describe a system with Fibonacci anyon excitations. Within 
this model, we apply existing general topological error-correction protocols, and 
simulate the successful preservation of quantum information encoded in topological 
degrees of freedom. Topological quantum computation protocols using nonabelian anyons 
typically implicitly assume the existence of an error-correction protocol to 
correct for diffusion or unwanted creation of anyons. Our results 
are the first explicit demonstration that such a scheme will 
be successful when applied to a universal topological quantum computer.
Although it is not surprising that the standard topological error-correction techniques we use successfully correct errors in universal nonabelian anyon systems, the ability to simulate the precise details of how and when these techniques succeed on finite system sizes has not previously been available. \cggb{Would a comment about the importance of locality in topological codes (compared to abstract QEC) maybe help Ref.~B's confusion about why we don't use ``conventional'' QEC? I'm not totally sure what he wants.}

@@

\cggb{I don't know how to address Ref.~B's comment here.}

@@

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the charge on 
each tile of the lattice, producing a \emph{syndrome} 
of occupied sites. Following this, it forms clusters of nearby occupied tiles and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when 
there is at most a single cluster remaining (see \Fref{f:decode}).

Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query the system by measuring these charges.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or 
when it performs a homologically nontrivial operation, thereby altering the 
codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that 
our qualitative results may be reproduced by most alternative families 
of decoders, although it may not be clear in all cases how to extend these decoders to the nonabelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering 
scheme is compatible with the structure in the noise 
processes that allows us to classically simulate them.

@@

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=\columnwidth]{anyons-kyle.pdf}
\caption{The decoder failure rate (a lower bound on the logical error probability) is shown as a function of simulation time for linear lattice sizes from $L=16$ to $128$. 
This provides strong evidence of threshold behavior at a critical memory lifetime of $t_{\mathrm{sim}}^*\simeq 0.125$. 
The existence of such a threshold implies that the Fibonacci anyon code simulated here is able to perfectly reliably store quantum information for times less than $t_{\mathrm{sim}}^*$ in the $L\to \infty$ limit. \cggb{Is this hedged enough? Maybe replace ``implies'' with ``would imply''?}}
\label{f:threshold}
\end{center}
\vspace{-10pt}
\end{figure}

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated. 
However, calculations of homologically nontrivial operations for small lattice sizes suggest that almost all such 
processes will indeed either cause a logical error or a complete failure to return to the codespace.
\cggb{Some kind of graph here to address Ref.~A's comment 3?}
As such we expect that neglecting the full simulation of 
these events does not significantly affect our observed threshold value.

@@

\section{Computation of homologically non-trivial operators}

>>>>

\newcommand{\Eref}[1]{Eq.~(\ref{#1})}
\newcommand{\Fref}[1]{Fig.~\ref{#1}}
\newcommand{\Aref}[1]{Appendix~\ref{#1}}

@@

Concretely, we consider the simulation of quantum error correction in a two-dimensional lattice 
system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum 
computation~\cite{Freedman2002, Nayak2008}. Fibonacci anyons are experimentally motivated as the 
expected excitations of the $\nu=\frac{12}{5}$ fractional quantum Hall 
states~\cite{Slingerland2001}, and can be realized in several spin 
models~\cite{Levin2005, Bonesteel2012, Kapit2013, Palumbo2014} and composite 
heterostructures~\cite{Mong2014}.
Any of these kinds of physical systems could be used to perform universal topological quantum computation \cggb{citations?}, and can be modelled by our simulations. Natural sources of noise from thermal fluctuations or external perturbations will be suppressed by the energy gap but must still be corrected to allow for scalable computation.

We use a flexible phenomenological model of dynamics and thermal 
noise to describe a system with Fibonacci anyon excitations. Within 
this model, we apply existing general topological error-correction protocols, and 
simulate the successful preservation of quantum information encoded in topological 
degrees of freedom. Topological quantum computation protocols using nonabelian anyons 
typically implicitly assume the existence of an error-correction protocol to 
correct for diffusion or unwanted creation of anyons. 
The ability to simulate the details of how and when these techniques succeed on finite system sizes has not previously been available, and so our results are the first explicit demonstration that such a scheme will be successful when applied to a universal topological quantum computer.

@@

@@

Our error-correction algorithm begins by measuring the charge on 
each tile of the lattice, producing a \emph{syndrome} 
of occupied sites. Following this, it joins nearby occupied tiles into clusters, and measures the total charge within each cluster.
Clusters with trivial charge are discarded, and then nearby clusters are merged (agglomeratively~\cite{Hastie2009}).
The merging process iterates at linearly increasing length scales, at each step measuring total charge and discarding trivial clusters, before concluding when 
there is at most a single cluster remaining (see \Fref{f:decode}).
This decoder is based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}.

Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so must repeatedly physically query the system to measure these charges.
This is also in contrast to abstract quantum codes, where syndrome extraction can typically be performed in one step.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or 
when it performs a homologically nontrivial operation, thereby altering the 
codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that 
our qualitative results may be reproduced by most alternative families 
of decoders, although it may not be clear in all cases how to extend these decoders to the nonabelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering 
scheme is compatible with the structure in the noise 
processes that allows us to classically simulate them.

@@

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=\columnwidth]{anyons-kyle.pdf}
\caption{The decoder failure rate (a lower bound on the logical error probability) is shown as a function of simulation time for linear lattice sizes from $L=16$ to $128$. 
This provides strong evidence of threshold behavior at a critical memory lifetime of $t_{\mathrm{sim}}^*\simeq 0.125$. 
The existence of such a threshold implies that the Fibonacci anyon code simulated here is able to perfectly reliably store quantum information for times less than $t_{\mathrm{sim}}^*$ in the $L\to \infty$ limit.}
\label{f:threshold}
\end{center}
\vspace{-10pt}
\end{figure}

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated.
However, naively computing the effect of the simplest possible homologically nontrivial operations consistently yields either a logical error or a complete failure to return to the codespace (see \Aref{s:homnontrivial}). As such it seems reasonable to suggest that neglecting the full simulation of 
these events does not significantly affect our observed threshold value.

@@

\section{Computation of homologically non-trivial operators}\label{s:homnontrivial}
